Proof by reflection for equalities in algebraic structures

### 2014-06-25 newPaper/paper.tex
changing name VariableA to Variable

bbb

@@

ccc

@@

ddd

\begin{code}[caption=Group, captionpos=b, label=lst1:haskell2]   
hasSymmetric : (c:Type) -> (p:dataTypes.Monoid c) -> c -> c -> Type
hasSymmetric c p a b = (Plus a b = Zero, Plus b a = the c Zero)

class dataTypes.Monoid c => dataTypes.Group c where
	Minus : c -> c ->
	Neg : c -> c
    Minus_simpl : (c1:c) -> (c2:c) -> Minus c1 c2 = Plus c1 (Neg c2) --Minus should not be primitive and should be simplifiable
    Plus_inverse : (c1:c) -> hasSymmetric c (%instance) c1 (Neg c1) -- Every element 'x' has a symmetric which is (Neg x)
\end{code}

@@

>>>>

A bit more interesting is the Semi-Group typeclass. A semi-Group is a Magma (ie, it still has a $plus$ operation), but it also has the property of associativity for this operation.

@@

The Monoid, for its part, is a Semi-Group with the property of neutral element.

@@

A Group is a Monoid, with two new operations. The binary $Minus$, and the unary $Neg$. One axiom says that the $Minus$ can always be simplified with the $Neg$ and the $Plus$. That means that $Minus$ is not a "primitive" operation of a Group.
The second axiom, which is the most important one for a Group, is the fact that any $c1$ admits $Neg c1$ as an inverse, where being an inverse if define just under.

\begin{code}[caption=Group, captionpos=b, label=lst1:haskell2]   
-- This is just a conjunctive predicate
hasSymmetric : (c:Type) -> (p:dataTypes.Monoid c) 
                -> c -> c -> Type
hasSymmetric c p a b = 
         (Plus a b = Zero, Plus b a = Zero)

class dataTypes.Monoid c => dataTypes.Group c where
	Minus : c -> c -> c
	Neg : c -> c
	Minus_simpl : (c1:c) -> (c2:c) -> 
	             Minus c1 c2 = Plus c1 (Neg c2) 
	Plus_inverse : (c1:c) -> 
	             hasSymmetric c (instance) c1 (Neg c1)
\end{code}

This hierarchy can be extended without difficulty with Abelian groups (with the axiom of commutativity), Ring, Commutative Rings...

The important thing to remember is that these typeclass will somehow be used as predicates on types. In order to call a solver on a type $T$, the user of the system will have to satisfy the corresponding typeclass by providing an instance of it for his type $T$, ie, he will have to prove that the corresponding properties effectively hold for $T$.

Note that the properties required by these typeclass for a specific type $T$ will either be proven if the operations on $T$ are real computable definitions, or axiomatised if the user is working in an axiomatised theory where the operations ($Plus$, $Neg$) are just axioms.

As discussed in the section above, the algorithm of normalization will not directly use the concrete value of type $T$, but a reflected term, indexed over the concrete value.

@@

We will have one datatype for reflected terms in each algebraic structure.

\begin{code}[caption=Reflected terms in a Magma, captionpos=b, label=lst1:haskell2]  
data ExprMa : Magma c -> (Vect n c) -> c -> Type where
    ConstMa : (p : Magma c) -> (g:Vect n c) -> (c1:c)  -> ExprMa p g c1 
    PlusMa : {p : Magma c} -> {g:Vect n c}  -> {c1:c} -> {c2:c} -> ExprMa p g c1 -> ExprMa p g c2 -> ExprMa p g (Plus c1 c2) 
    VarMa : (p:Magma c) -> {g:Vect n c} -> {c1:c} -> VariableA (magma_eq_as_elem_of_set p) g c1 -> ExprMa p g c1
\end{code}
